---
title: "Data Viz FA 2020: Week 7 Assignment"
author: "Ahmed Khan"
date: "10/17/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(dplyr)

setwd("/Users/ahmedkhan/Downloads")

NILS = read_csv("WHO_NILS_ALL.csv")
FAO = read_csv("UNFAO_by_Year.csv")
GFSI = read_csv("GlobalFoodSecurityIndex_2019.csv")

#Long format the FAO data (prep for Tidyverse use)
FAO = FAO %>% pivot_longer(cols = starts_with("20"), names_to = "year", values_to = "values")

#Fix year format
FAO$year = as.numeric(FAO$year)

#Format the NILS data, select certain variables to plot
NILS_Tibble = as.tibble(NILS)

NILS_Tibble$Indicator =  
  recode(NILS_Tibble$Indicator, 
         "Population below minimum level of dietary energy requirement (undernourishment) (%)" = "% Undernourished",
         "Seats held by women in national parliament (%)" = "% Women in Parliament",
         "Averaged aggregate governance indicators" = "Avg Governance Indicator",
         "Total expenditure on health as % of gross domestic product" = "Health Spend %",
         "Gender Inequality Index (GII)" = "Gender Inequality Index",
         "Global Hunger Index (GHI)" = "Global Hunger Index"
         )

#Exploring variables
#Exploring variables
parl_seats = filter(NILS_Tibble, Indicator =="% Women in Parliament")
  #Get latest value available
  parl_seats_max = parl_seats %>% group_by(Country) %>%
    slice(which.max(Year)) %>%
    select(Country, Value)

Undnourish = filter(NILS_Tibble, Indicator =="% Undernourished")
  #Has dups and no year info so just get best amount (in this case lowest under-nourished %) and subtract from 100% to get % Nourished 
  Undnourish_Max = Undnourish %>% group_by(Country) %>%
    summarize(nourished = 100-min(Value))

Health_spend = filter(NILS_Tibble, Indicator =="Health Spend %")
  #Get latest value available
  Health_Max = Health_spend %>% group_by(Country) %>%
    slice(which.max(Year)) %>%
    select(Country, Value) %>%
    rename(healthspend = Value)

Governance = filter(NILS_Tibble, Indicator =="Avg Governance Indicator")
  #Get latest score available
  Governance_Max = Governance %>% group_by(Country) %>%
    slice(which.max(Year)) %>%
    select(Country, Value) %>%
    rename(governance = Value)

  
#Join the data
part_seats_nrsh = 
  parl_seats_max %>%
  left_join(Undnourish_Max, by = "Country") %>%
  left_join(Health_Max, by = "Country") %>%
  left_join(Governance_Max, by = "Country") %>%
  drop_na(nourished)

```

## Summary Narrative

This report shows some high level metrics of dietary indicators.

Figure 1 shows that, on average, as the percent of women in parliament increases there is not a real association with improved rates of nourishment in the population. However, as the colors of the points indicate, countries with better governance scores tend to have a much higher percent of the population that meets minimum nourishment needs. Governance is measured as the extent to which governments score well on accountability to citizens, effectively responding to citizen needs, control of corruption, political stability, regulatory environment management, and maintaining the rule of law.
Note: This data excludes about 80 coountries for which Nutrition measures were not available.

```{r example}
#summary(cars)
```


```{r pressure, echo=FALSE, message=FALSE, warning = FALSE}
#Country Level using NILS Data

#Plot relationship between % of women in parliament and undernourishment
#No real relationship found b/w women and nourishment
#However, colors suggest that beter governance = higher nourishment rates :)
ggplot(part_seats_nrsh, aes(x=Value, y=nourished, color = governance)) +
  geom_jitter() + 
  geom_smooth(se=FALSE, method = "loess", size = 0.4, color ="gray")+
  theme_classic() +
  ylim(50,110)+
  theme(plot.title = element_text(face="bold")) +
  labs(x="% Women in Parliament",y="% Nourished Population ",title="Fig 1. Sufficiently Nourished Population by Country", subtitle = "Data points are countries (using latest available indicators)", caption = "Higher Governance Score is better. Source: World Health Organization NLiS", color = "Governance Score")+
scale_color_gradient2(low="red", high="green")
#+scale_color_gradientn(colours=topo.colors(3),na.value = "transparent",
    #                    n.breaks=3,labels=c("Worst","Neutral","Best"))
```
